---
title: "Assignment 13"
author: "Stacey Schwarcz"
date: "Sunday, November 30, 2014"
output: html_document
---

Question 3 Instructions:
Find an R package that covers the same functionality as some base R code. Write some code to benchmark probably with loaded or generated data of a reasonable size--to compare speed in the two environments). Publish your code and benchmark results at rpubs.com. 

Example: Compare the speed of a dplyr aggregation function and the corresponding base R apply() function against a reasonably-sized dataset; if you want, you may instead compare the speed of two R packages that perform the same function.

Import the Jeopardy questions data set
source: http://j-archive.com/

```{r}
jeopardy<-read.csv("C:/temp/JEOPARDY_CSV.csv", header = TRUE, sep = ",")
head(jeopardy)
#convert data in Value column to numeric
jeopardy$Value=as.character(jeopardy$Value)
jeopardy$Value=sub(",","",jeopardy$Value)
jeopardy$Value=sub("\\$","",jeopardy$Value)
jeopardy$Value=as.numeric(jeopardy$Value)
options(digits=0)

```
Find the average dollar value for questions by category. 

Calculate using the base R aggregate function:
```{r}
agg<-setNames(aggregate(jeopardy$Value, list(category = jeopardy$Category), mean),c('Category','meanValue'))
head(agg)
```

Do the same calculation with ddpy in the plyr package:
```{r}
library(plyr)
dd<-ddply(jeopardy, .(Category), summarise, meanValue = mean (Value))
head(dd)
```
Time to execute with aggregrate and then ddply:
```{r}
system.time(aggregate(jeopardy$Value, list(category = jeopardy$Category), mean))
system.time(ddply(jeopardy, .(Category), summarise, meanValue = mean (Value)))

```
ddply take 3 times as long as aggregate for this calculation

link to published file:
http://rpubs.com/sms2015/45998
